---
title: "data manipulation"
author: "Xing Chen"
date: 2019-9-19
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```


## import data files

```{r}
litters_data = read_csv("./data/FAS_litters.csv")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv")
pups_data = janitor::clean_names(pups_data)
```


## selecting...

```{r}
select(litters_data, group, litter_number, starts_with("pups"))

select(litters_data, litter_number, group)

select(litters_data, litter_number, group, everything())

select(litters_data, -group)

select(litters_data, litter_number, gd0_weight:pups_born_alive)

litters_select = select(litters_data, litter_number, gd0_weight:pups_born_alive)

select(litters_data, litter_number, GROUP = group)

rename(litters_data, Group = group)

litters_data
```

In larger datasets, I also use `starts_with()`, `ends_with()`, and `contains()` often.


## filtering.....

```{r, results='hide'}
filter(litters_data, group == "Con7")

filter(litters_data, gd_of_birth == 20)

filter(litters_data, group %in% c("Con7", "Con8"))

filter(litters_data, gd0_weight + gd18_weight < 70)

# dont do this!
# filter(litters_data, !is.na(gd0_weight))

drop_na(litters_data)
drop_na(litters_data, gd0_weight)
```


## mutate...

```{r}
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group),
  group = str_to_upper(group)
)
```

## Arrange...

```{r}
arrange(litters_data, pups_born_alive)

arrange(litters_data, desc(pups_born_alive))

arrange(litters_data, pups_born_alive, gd0_weight)

```


## pipe....

create a collection of commands...

```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)
  ) %>% 
  drop_na(gd0_weight)
```

By default, the pipe will take the result of one function call and use that as 
the first argument of the next function call

when the result of one function call is not the first argument of the next 
call, that is the point you have to worry and think about. But dont worry
too much about this.

```{r, eval=FALSE}
litters_data %>% view()

litters_data %>% pull(gd0_weight) %>% mean()
```







